---
title: "Time series numbers"
author: "Adrià Auladell"
date: "`r format(Sys.time(), '%d %B, %Y, %H:%M')`"
knit: (function(inputFile, encoding) { 
      out_dir <- 'results/tests/';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(here::here(),
                        out_dir, 'seasonality.html')) })
output:
  html_document:
    number_sections: yes
    toc: yes
---

# Packages 
```{r}
knitr::opts_chunk$set(fig.width = 12, fig.height = 9)
```

```{r}
library(tidyverse)
library(speedyseq)
library(here)
```

```{r}
source(here("src/util/backbone_functions.R") )
source(here("src/util/backbone_params-graphs.R") )
source(here("src/util/main_functions.R"))
source(here("src/util/timeseries_functions.R"))

```


# Data import
```{r}
bl.phy <- readRDS(here('data/cleaned/blphyloseq.rds') )
bl.phy.filt <- readRDS(here('data/cleaned/blphy_filt_7sams.rds') )
lombsea <- readRDS(here('data/analysis/lombsea.rds'))
abunprev <- readRDS(here('data/analysis/abund_prev.rds'))

tax <- as(tax_table(bl.phy.filt), 'matrix') %>% as_tibble(rownames = 'asv')
```

```{r}
  
geoMean = function(x, na.rm=TRUE){
  exp(sum(log(x[x > 0]), na.rm=na.rm) / length(x))
}

bl.phy.clr <- bl.phy %>% 
  transform_sample_counts(., function(x) x+1) %>%
  transform_sample_counts(function(x) log(x/geoMean(x)))
```

# Distribution of behaviours 

```{r}
abprevtax <- abunprev %>% 
  left_join(tax, by = 'asv') %>% 
  mutate(seasonal = ifelse(asv %in% lombsea$asv, TRUE, FALSE)) 
  
abprevsea <- abprevtax %>% 
  filter(asv %in% lombsea$asv) %>% 
  select(asv, behavior, presence)
  

abprevsea %>% 
  select(behavior, presence) %>% 
  table()

abprevsea %>% filter(behavior == 'Broad') 
abunprev %>% filter(behavior == 'Broad') %>% 
  left_join(tax, by = 'asv')


abprevtax %>% filter(behavior == 'Narrow') %>% pull(class) %>% table()
```
# Plotting specific winter 

```{r}

# bl.phy.asinh <- transform_sample_counts(bl.phy.filt, function(x) asinh(x))


num.days.mnt <- c(0,31,28,31,30,31,30,31,31,30,31,30)
cumnum <- cumsum(num.days.mnt)

sea.asvs <- psmelt(bl.phy.clr) %>% 
  filter(OTU %in% c('asv122', 'asv55', 'asv131')) %>% 
  ggplot(aes(day_of_year, Abundance)) + 
  geom_jitter(aes(color = OTU), alpha = 0.6, show.legend = F) + 
  stat_smooth(aes(x = day_of_year,
                  group = OTU,
                  color = OTU), # continuous x-axis
              method = "gam",
              formula = y ~ s(x, k =12, bs = 'cc'),
              se = F, size = 2, show.legend = F) + 
  facet_wrap(~str_c(class, ', ', str_to_upper(OTU))) + 
  lil.strip + 
  ylab('Centered log ratio (log10( ASV read count / geoMean))') + 
  scale_x_continuous(breaks = cumnum,
                     name = 'Month',
                     labels = str_to_title(month.order)) +
  scale_color_solarized(name = 'ASV')

ggsave(filename = here('results/figures/timeseries/verymuch_seasonal_asvsclr.pdf'),
       plot = sea.asvs, 
       width = 11, 
       height = 5)


this <- psmelt(bl.phy.clr) %>% 
  filter(OTU %in% c('asv122', 'asv55', 'asv131'))


this %>% select(OTU, phylum, class, seq) %>% distinct()

this %>% pull(seq) %>% as.character() %>%  unique()
```

# Comparsion trends ASVs
```{r}
vals <- abprevtax %>% 
  pull(genus) %>% 
  table() %>% sort() 

vals 


```
How do they compare, Pelagibacter and HIMB59 ASVs
```{r}

genera <- vals %>% 
           names() %>% tail( n =6)

asvs <- abprevtax %>% 
  filter(genus %in% c(genera, 'Synechococcus_C', 'UBA7446')) %>% 
  pull(asv)

psm.asinh <-  psmelt(transform_sample_counts(bl.phy.filt, function(x) asinh(x)))


psm.asinh %>% 
  filter(OTU %in% asvs) %>% 
  ggplot(aes(month, Abundance)) + 
  geom_jitter(alpha = 0.6) + 
  stat_smooth(aes(x = month,
                  group = OTU,
                  color = family),
              method = "gam",
              formula = y ~ s(x, k =12, bs = 'cc'),
              se = F, size = 2,
              show.legend = F, alpha = 0.7) + 
  facet_wrap(~genus)
```

```{r}
pelagsvs <- abprevtax %>% 
  filter(order %in% c('Pelagibacterales', 'HIMB59')) %>% 
  pull(asv)

noseacladeII <- tax %>% 
  filter(genus == 'Pelagibacter_A') %>% 
  pull(asv)

psm.asinh %>% 
  filter(OTU %in% c(pelagsvs, noseacladeII)) %>% 
  ggplot(aes(month, Abundance)) + 
  geom_jitter(alpha = 0.6) + 
  stat_smooth(aes(x = month,
                  group = OTU,
                  color = genus),
              method = "gam",
              formula = y ~ s(x, k =12, bs = 'cc'),
              se = F, size = 2,
              show.legend = F, alpha = 0.7) + 
  facet_wrap(~genus)


pelagoverview <- tax %>% 
  filter(genus %in% c('Pelagibacter', 'Pelagibacter_A', 'AG-337-102',
                      'HIMB59', 'HIMB114')| order %in% 'HIMB59' ) %>% 
  pull(asv)

psm.asinh %>% 
  filter(OTU %in% pelagoverview) %>% 
  mutate(seasonal = ifelse(OTU %in% lombsea$asv, TRUE, FALSE)) %>% 
  ggplot(aes(month, Abundance)) + 
  geom_jitter(alpha = 0.6) + 
  stat_smooth(aes(x = month,
                  group = OTU,
                  color = genus),
              method = "gam",
              formula = y ~ s(x, k =12, bs = 'cc'),
              se = F, size = 2,
              show.legend = F, alpha = 0.7) + 
  facet_wrap(~genus + seasonal)

```

Let's see Cyanobac and Flavobacteria 

```{r}

cyanoflavo <- tax %>% 
  filter(family %in% c('Cyanobiaceae' )) %>% 
  pull(asv)

psm.asinh %>% 
  filter(OTU %in% cyanoflavo) %>% 
  mutate(seasonal = ifelse(OTU %in% lombsea$asv, TRUE, FALSE)) %>% 
  ggplot(aes(month, Abundance)) + 
  geom_jitter(alpha = 0.6) + 
  stat_smooth(aes(x = month,
                  group = OTU,
                  color = genus),
              method = "gam",
              formula = y ~ s(x, k =12, bs = 'cc'),
              se = F, size = 2,
              show.legend = F, alpha = 0.7) + 
  facet_wrap(~genus + seasonal)

# only flavo

cyanoflavo <- tax %>% 
  filter(family %in% c('Flavobacteriaceae')) %>% 
  pull(asv)

psm.asinh %>% 
  filter(OTU %in% cyanoflavo) %>% 
  mutate(seasonal = ifelse(OTU %in% lombsea$asv, TRUE, FALSE)) %>% 
  ggplot(aes(month, Abundance)) + 
  geom_jitter(alpha = 0.6) + 
  stat_smooth(aes(x = month,
                  group = OTU,
                  color = genus),
              method = "gam",
              formula = y ~ s(x, k =12, bs = 'cc'),
              se = F, size = 2,
              show.legend = F, alpha = 0.7) + 
  facet_wrap(~genus + seasonal)

```

What about SAR86

```{r}
otus.decipher <- readRDS(here('data/analysis/bbmo_otus_decipher.rds')) 

otu99 <- data.frame( OTU = otus.decipher$asv,
                     otu99 = pull(otus.decipher, cluster99) )


plot_clusters_asvs <- function(selasvs){
  
  psm.asinh %>% 
    filter(OTU %in% selasvs) %>% 
    left_join(otu99, by = 'OTU') %>% 
    mutate(seasonal = ifelse(OTU %in% lombsea$asv, TRUE, FALSE)) %>% 
    mutate(labol = ifelse(month == '12' & year == 2013, OTU, NA_character_)) %>% 
    ggplot(aes(month, Abundance)) + 
    stat_smooth(aes(x = month,
                    group = OTU,
                    color = as.character(seasonal)),
                method = "gam",
                formula = y ~ s(x, k =12, bs = 'cc'),
                se = F, size = 1,
                show.legend = F) + 
    ggrepel::geom_label_repel(aes(label = labol,
                                  color = as.character(seasonal)),
                              size = 2, na.rm = T)+ 
    scale_color_manual(values = c('black', 'red'), name = 'Seasonal') +
    lil.strip + 
    leg.bottom + 
    scale_x_month + 
    facet_wrap(~cluster)
  
}


```

```{r}

sar86asvs <- tax %>% 
  filter(genus %in% c('SAR86A')) %>% 
  pull(asv)

plot_clusters_asvs(sar86asvs)
```


```{r}

rhodobacasvs <- tax %>% 
  filter(family %in% c('Rhodobacteraceae')) %>% 
  pull(asv)

plot_clusters_asvs(rhodobacasvs)

```

```{r}
d472 <- tax %>% 
  filter(family %in% c('D2472')) %>% 
  pull(asv)

plot_clusters_asvs(d472)

```
```{r}

puni <- tax %>% 
  filter(order %in% c('Puniceispirillales')) %>% 
  pull(asv)

plot_clusters_asvs(puni)
```

```{r}
himbi <- tax %>% 
  filter(order %in% c('HIMB59')) %>% 
  pull(asv)

plot_clusters_asvs(himbi)
```


```{r}

flav <- tax %>% 
  filter(family %in% c('Flavobacteriaceae')) %>% 
  pull(asv)


plot_clusters_asvs(flav)
```

```{r}
pelaasvs <- tax %>% 
  filter(genus %in% c('Pelagibacter')) %>% 
  pull(asv)


plot_clusters_asvs(pelaasvs)
```


```{r}
aegeanasvs <- tax %>% 
  filter(genus %in% c('AG-337-I02')) %>% 
  pull(asv)

plot_clusters_asvs(aegeanasvs)
```

```{r}

lumiasvs <- tax %>% 
  filter(genus %in% c('Luminiphilus')) %>% 
  pull(asv)

plot_clusters_asvs(lumiasvs)
```
```{r}

pelagiaaasvs <- tax %>% 
  filter(genus %in% c('Pelagibacter_A')) %>% 
  pull(asv)

plot_clusters_asvs(pelagiaaasvs)
```

```{r}
syneasvs <- tax %>% 
  filter(genus %in% c('Synechococcus_C')) %>% 
  pull(asv)

plot_clusters_asvs(syneasvs)


```

# Contribution to total relative abundance 

```{r}

lomb.02 <- readRDS(here('data/analysis/lomball.rds'))

# seeing how does this change with these values 
lomb.sea <- tibble( asv = names(lomb.02),
                    pval = map_dbl(lomb.02, ~.x$p.value),
                    peak = map_dbl(lomb.02, ~.x$peak),
                    interval  = map(lomb.02, ~.x$peak.at),
                    int.min = map_dbl(interval, ~.[[2]]),
                    int.max = map_dbl(interval, ~.[[1]])) %>% 
  mutate( qval = fdrtool::fdrtool(pval, statistic = 'pvalue')$qval) %>% 
  filter(qval <= 0.01, peak >= 10, int.max <= 2)


season.relab <- taxa_sums(bl.phy)[lombsea$asv] %>% sum()
total.relab <- taxa_sums(bl.phy) %>% sum()
total.tested.relab <- taxa_sums(bl.phy.filt) %>% sum()

season.relab * 100 / total.relab
season.relab * 100 / total.tested.relab


sam.sea <- otu_table(bl.phy)[,lombsea$asv] %>% rowSums()
total.sea <- otu_table(bl.phy) %>% rowSums()

relabs <- sam.sea / total.sea 

mean(relabs)
sd(relabs)

therel <-  data.frame(this  = relabs, 
           sample_data(bl.phy))

therel %>% 
  ggplot( aes(this)) + 
  geom_histogram()


ggplot(therel, 
       aes(month, this)) + 
  geom_jitter(alpha = 0.7)
      
```

## Contrib for each behav

```{r}

sum(taxa_sums(bl.phy)[abprevsea %>% filter(behavior == 'Broad') %>% pull(asv)]) * 100 / total.relab
sum(taxa_sums(bl.phy)[abprevsea %>% filter(behavior == 'Narrow') %>% pull(asv)]) * 100 / total.relab
sum(taxa_sums(bl.phy)[abprevsea %>% filter(behavior %in% c('Other', 'CRT')) %>% pull(asv)]) * 100 / total.relab
 
```

# Taking the genus as a unit, which one presents a higher seasonality? 

```{r}
lombsea %>% 
  left_join(tax, by = 'asv') %>% 
  group_by(genus) %>% 
  filter(n() >=3) %>%
  summarize(count = n(),
            mean.peak = mean(peak)) %>% 
  arrange(-mean.peak) 


lombsea %>% 
  left_join(tax, by = 'asv') %>% 
  arrange(-peak)

```

Some of the ASVs present a peak of 42. 
This is surprising, let's see how is the distribution 


```{r}
plot_monthts(bl.phy.filt, asvs = 'asv122')
plot_monthts(bl.phy.filt, asvs = 'asv55')
plot_monthts(bl.phy.filt, asvs = 'asv25')
plot_monthts(bl.phy.filt, asvs = 'asv29')
plot_monthts(bl.phy.filt, asvs = 'asv39')
```


